Statistical performance analysis of neural mass models?

### 2013-06-13 manuscript/manuscript.tex
a
minor change
added plot for example data

\section{Background}

@@

The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$). The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns. Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}. Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
\begin{figure}[ht]
	\centering
		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz.pdf}
	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx. = Cortex, Th. = Thalamus). The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.  a) Minimal model of a column with inhibitory feedback \cite{Silva1974}. b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}. c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
	\label{fig:NMMs}
\end{figure}
The parameters of the neural masses not only define the population type, but also the behaviour exhibited by the model. For example, certain parameter combinations result in a model of a cortical column that will generate alpha wave type activity (normal activity) and, for another set of parameters, we create a model that will exhibit epileptic behaviour \cite{Wendling2002}. Therefore, we consider a family of models, which we define generally as 
\begin{align}\label{eq:NeuralMassModel}
    \dot{\mathbf{x}}(t) =& f_{\theta}\left(\mathbf{x}(t),\mathbf{u}(t)\right)\\
    y(t) =& \mathbf{C}\mathbf{x}(t) + e(t),
\end{align}
where $\mathbf{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\mathbf{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies. The function $f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits. The iEEG is denoted by $y(t)$, $\mathbf{C}$ is the observation matrix, and $e(t)$ is the observation noise.

>>>>

Not claiming anything about which model is the 'best', but simply what level of complexity is possible - how increasing complexity affects uncertainty.
\section{Background}

@@

The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$). The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns. Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}. Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
\begin{figure}[ht]
	\centering
		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz.pdf}
	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx. = Cortex, Th. = Thalamus). The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.  a) Minimal model of a column with inhibitory feedback \cite{Silva1974}. b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}. c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
	\label{fig:NMMs}
\end{figure}

\begin{figure}[ht]
	\centering
		\includegraphics[scale=1]{./figures/pdf/Example_Data.pdf}
		\caption{\textbf{Example data generated using the neural mass models.}}
	\label{fig:NMMs}
\end{figure}

The parameters of the neural masses not only define the population type, but also the behaviour exhibited by the model. For example, certain parameter combinations result in a model of a cortical column that will generate alpha wave type activity (normal activity) and, for another set of parameters, we create a model that will exhibit epileptic behaviour \cite{Wendling2002}. Therefore, we consider a family of models, which we define generally as 
\begin{align}\label{eq:NeuralMassModel}
    \dot{\mathbf{x}}(t) =& f_{\theta}\left(\mathbf{x}(t),\mathbf{u}(t)\right)\\
    y(t) =& \mathbf{C}\mathbf{x}(t) + e(t),
\end{align}
where $\mathbf{x}(t) \in \mathbb{R}^{n_x}$ is a state vector representing the postsynaptic membrane potentials generated by each population synapse and their time derivatives, $n_x$ is the number of states and $\mathbf{u}(t)$ represents the system input, which may be from afferent connections or other brain regions, exogenous input, or account for model inaccuracies. The function $f_{\theta}(\cdot)$ describes the dynamics, where $\theta \in \mathbb{R}^{n_{\theta}}$ determines the model type and the behaviour it exhibits. The iEEG is denoted by $y(t)$, $\mathbf{C}$ is the observation matrix, and $e(t)$ is the observation noise.
